---
title: "COVID-19 Sentinel Surveillance"
author: "AMPATH Population Health | Busia County"
output: 
  rmdformats::material:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango

---

```{r setup, include=FALSE}
options(java.parameters = "-Xmx15g")

knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      echo=FALSE,
                      #dpi=96,
                     # fig.width=7,# fig.height=4, # Default figure widths
                     # dev="png", #dev.args=list(type="cairo"), # The png device
                      # Change to dev="postscript" if you want the EPS-files
                      # for submitting. Also remove the dev.args() as the postscript
                      # doesn't accept the type="cairo" argument.
                      error=FALSE)
 
# Evaluate the figure caption after the plot
#knitr::opts_knit$set(eval.after='fig.cap')
 
# Use the table counter that the htmlTable() provides
options(table_counter = TRUE)
 
# Use the figCapNo() with roman letters
#options(fig_caption_no_roman = TRUE)
#options(kableExtra.latex.load_packages = F)

# Then install the Grmd-package by running below code:
#devtools::install_github("gforge/Grmd")


# function to install missing packages
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, repos='http://cran.rstudio.com/')
  sapply(pkg, require, character.only = TRUE)
}

#install.packages('package_name', dependencies=TRUE, repos='http://cran.rstudio.com/')

packages =c( "dplyr",  "readxl","Hmisc","Gmisc", "magrittr", "flextable", "MASS", "tidyverse", "caret", "knitr", "kableExtra","xtable", "stargazer", "ggpubr", "haven", "PerformanceAnalytics", "naniar", "gridExtra", "ggthemes", "TSstudio")

ipak(packages)

# packages not pubished
#devtools::install_github("kassambara/easyGgplot2")
library(easyGgplot2)

# packages not pubished
#devtools::install_github('cttobin/ggthemr')
#library(ggthemr)


select = dplyr::select; summarize = dplyr::summarize; rename = dplyr::rename; mutate = dplyr::mutate;

#ggthemr("flat")

#source("unbalanced_functions.R")
source("table1_utils.R")
```

# Explaratory Data Analysis | Tables

The goal of the exploratory analysis is to identify the key predictors, generate insights and dicover trends and patters. Throughout this process we are mainly interested in detecting:

* Correlation and assoicitations
* Trends in time e.g viral load
* Outliers
* Any other key information that might be useful in ML modeling




```{r warning=F,cache=T}
# Import Dataset
source("import.r")
dataset=data

#unique(dataset$site_name.factor)

#str(dataset)
```



```{r fig.align='center',warning=FALSE, cache=F}
na.rm.var=T;stats=T
dataset=data
unique(dataset$gender.factor)
# Getting descriptive statistics 
mergeDesc(
  
  "Age" = getTable1Stats("age","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),  
  "Gender" = getTable1Stats("gender.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Marital Status" = getTable1Stats("marital_status.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Eduaction Level" = getTable1Stats("what_is_your_educational.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Occupation" = getTable1Stats("what_is_your_occupation.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var)
          
          )%>%
   htmlTable( caption  = "<b> Socio-demographics </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1,4,1 ), cgroup = c('', 'Sentinel Sites', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 




mergeDesc(
  
  "Travelled outside the county" = getTable1Stats("have_you_travelled_outside.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),  
   "Previous tested" = getTable1Stats("have_you_been_tested_befor.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
   "Results of previous test" = getTable1Stats("what_was_the_test_result.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Breathing / respiratory symptoms" = getTable1Stats("have_you_experienced_any_b.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Known comorbidities" = getTable1Stats("do_you_have_any_comorbidit.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Exposed to known +ve Contacts" = getTable1Stats("have_you_had_a_contact_wit.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Severity" = getTable1Stats("severity_information.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Patients recovered" = getTable1Stats("patient_recovered.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Patients hospitalized" = getTable1Stats("patient_hospitalized.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Covid19 test results" = getTable1Stats("covid_result.factor","site_name.factor",desc_both,statistics=stats,na.rm=na.rm.var)
          
          )%>%
   htmlTable( caption  = "<b> Clinicals / Syndromic Summaries </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1,4,1 ), cgroup = c('', 'Sentinel Sites', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 




mergeDesc(
  
  # 
  #  "Age" = getTable1Stats("age","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),  
  # "Gender" = getTable1Stats("gender.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  # "Marital Status" = getTable1Stats("marital_status.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  # "Eduaction Level" = getTable1Stats("what_is_your_educational.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
   "Occupation" = getTable1Stats("what_is_your_occupation.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
   "Sentinel Sites" = getTable1Stats("site_name.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Travelled outside the county" = getTable1Stats("have_you_travelled_outside.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),  
   "Previous tested" = getTable1Stats("have_you_been_tested_befor.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
   "Results of previous test" = getTable1Stats("what_was_the_test_result.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Breathing / respiratory symptoms" = getTable1Stats("have_you_experienced_any_b.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Known comorbidities" = getTable1Stats("do_you_have_any_comorbidit.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Exposed to known +ve Contacts" = getTable1Stats("have_you_had_a_contact_wit.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Severity" = getTable1Stats("severity_information.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Patients recovered" = getTable1Stats("patient_recovered.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var),
  "Patients hospitalized" = getTable1Stats("patient_hospitalized.factor","covid_result.factor",desc_both,statistics=stats,na.rm=na.rm.var)
 
          
          )%>%
   htmlTable( caption  = "<b> Positivity Association </b>",useViewer=T,ctable = TRUE, align = 'lcccc',
              n.cgroup = c(1,2,1 ), cgroup = c('', 'Covid19 Test Results', '') ,tfoot="<sup>&Dagger;</sup>"
              )%>% htmlTable::addHtmlTableStyle(col.rgroup = c("#FFF", "#EEF")) 


```

# Explaratory Data Analysis | Univariate Plots

For each continuous column, visually check the following: 

1. Variation in the column
2. Its distribution
3. Any outliers
4. q-q plot with normal distribution


```{r fig.align='center', cache=T, fig.width=5, warning=FALSE, eval=T}
suppressMessages({library(highcharter)})
# Reverse side
dataset %>%
  count(site_name.factor,what_is_your_occupation.factor)%>%
  na.omit()%>% 
  arrange(desc(n),site_name.factor) %>% 
hchart( "column", hcaes(x = site_name.factor, y = n, group = what_is_your_occupation.factor),
        options3d = list(enabled = TRUE, beta = 0, alpha = 0))%>%
  hc_plotOptions(series = list(dataLabels = list(enabled = TRUE))) %>%
 hc_plotOptions(series = list(#stacking = "normal",  
                               dataLabels = list(enabled = TRUE))) %>%
  hc_xAxis(title = list(text = "")) %>% 
   hc_yAxis(title = list(text = "Counts")) %>% 
  hc_chart( options3d = list(enabled = TRUE, beta = 0, alpha = 0))


dd =dataset%>% 
  select(site_name.factor) %>% 
  filter(!is.na(site_name.factor)) %>% 
  count(site_name.factor) %>%
  mutate(name=site_name.factor)
  
  highchart() %>% hc_add_series(dd,"pie",innerSize="50%",hcaes(name, n)) %>% 
   hc_plotOptions(pie =list(dataLabels = list(enabled = TRUE,format="{point.name}: {point.y} ({point.percentage:.2f}%)")))%>%
   hc_legend(align = "left", verticalAlign = "top",
            layout = "vertical", x = 0, y = 100) 

```




```{r fig.align='center', cache=T, fig.width=5, warning=FALSE, eval=T}
dd =dataset%>% 
  select(covid_result.factor) %>% 
  filter(!is.na(covid_result.factor)) %>% 
  count(covid_result.factor) %>%
  mutate(name=covid_result.factor)
  
  highchart() %>% hc_add_series(dd,"pie",innerSize="50%",hcaes(name, n)) %>% 
   hc_plotOptions(pie =list(dataLabels = list(enabled = TRUE,format="{point.name}: {point.y} ({point.percentage:.2f}%)")))%>%
   hc_legend(align = "left", verticalAlign = "top",
            layout = "vertical", x = 0, y = 100)  %>%  hc_title(text = "Overall")
  

dd =dataset%>% 
  select(covid_result.factor,site_name.factor) %>% 
    filter(site_name.factor=="KOCHOLYA") %>% 
   filter(!is.na(covid_result.factor)) %>% 
  count(covid_result.factor) %>%
  mutate(name=covid_result.factor)
  
  highchart() %>% hc_add_series(dd,"pie",innerSize="50%",hcaes(name, n)) %>% 
   hc_plotOptions(pie =list(dataLabels = list(enabled = TRUE,format="{point.name}: {point.y} ({point.percentage:.2f}%)")))%>%
   hc_legend(align = "left", verticalAlign = "top",
            layout = "vertical", x = 0, y = 100)  %>%  hc_title(text = "KOCHOLYA")
  
  
  dd =dataset%>% 
  select(covid_result.factor,site_name.factor) %>% 
      filter(site_name.factor=="NAMBALE") %>% 
  filter(!is.na(covid_result.factor)) %>% 
  count(covid_result.factor) %>%
  mutate(name=covid_result.factor)
  
  highchart() %>% hc_add_series(dd,"pie",innerSize="50%",hcaes(name, n)) %>% 
   hc_plotOptions(pie =list(dataLabels = list(enabled = TRUE,format="{point.name}: {point.y} ({point.percentage:.2f}%)")))%>%
   hc_legend(align = "left", verticalAlign = "top",
            layout = "vertical", x = 0, y = 100)  %>%  hc_title(text = "NAMBALE")
  
   dd =dataset%>% 
  select(covid_result.factor,site_name.factor) %>% 
      filter(site_name.factor=="MATAYOS") %>% 
  filter(!is.na(covid_result.factor)) %>% 
  count(covid_result.factor) %>%
  mutate(name=covid_result.factor)
  
  highchart() %>% hc_add_series(dd,"pie",innerSize="50%",hcaes(name, n)) %>% 
   hc_plotOptions(pie =list(dataLabels = list(enabled = TRUE,format="{point.name}: {point.y} ({point.percentage:.2f}%)")))%>%
   hc_legend(align = "left", verticalAlign = "top",
            layout = "vertical", x = 0, y = 100)  %>%  hc_title(text = "MATAYOS")


```


```{r fig.align='center', cache=T, fig.width=5, warning=FALSE, eval=T}

# Reverse side
dataset %>%
  count(what_is_your_occupation.factor,covid_result.factor)%>%
  na.omit()%>% 
  arrange(desc(n),what_is_your_occupation.factor) %>% 
hchart( "column", hcaes(x = what_is_your_occupation.factor, y = n, group = covid_result.factor),
        options3d = list(enabled = TRUE, beta = 0, alpha = 0))%>%
  hc_plotOptions(series = list(dataLabels = list(enabled = TRUE))) %>%
 hc_plotOptions(series = list(#stacking = "normal",  
                               dataLabels = list(enabled = TRUE))) %>%
  hc_xAxis(title = list(text = "")) %>% 
   hc_yAxis(title = list(text = "Counts")) %>% 
  hc_chart( options3d = list(enabled = TRUE, beta = 0, alpha = 0)) %>%  hc_title(text = "Occupation Positivity")


# Reverse side
dataset %>%
  count(have_you_experienced_any_b.factor,covid_result.factor)%>%
  na.omit()%>% 
  arrange(desc(n),have_you_experienced_any_b.factor) %>% 
hchart( "column", hcaes(x = have_you_experienced_any_b.factor, y = n, group = covid_result.factor),
        options3d = list(enabled = TRUE, beta = 0, alpha = 0))%>%
  hc_plotOptions(series = list(dataLabels = list(enabled = TRUE))) %>%
 hc_plotOptions(series = list(#stacking = "normal",  
                               dataLabels = list(enabled = TRUE))) %>%
  hc_xAxis(title = list(text = "")) %>% 
   hc_yAxis(title = list(text = "Counts")) %>% 
  hc_chart( options3d = list(enabled = TRUE, beta = 0, alpha = 0)) %>%  hc_title(text = "Breathing / Respiratory Symptoms Positivity")

# Reverse side
dataset %>%
  count(severity_information.factor,covid_result.factor)%>%
  na.omit()%>% 
  arrange(desc(n),severity_information.factor) %>% 
hchart( "column", hcaes(x = severity_information.factor, y = n, group = covid_result.factor),
        options3d = list(enabled = TRUE, beta = 0, alpha = 0))%>%
  hc_plotOptions(series = list(dataLabels = list(enabled = TRUE))) %>%
 hc_plotOptions(series = list(#stacking = "normal",  
                               dataLabels = list(enabled = TRUE))) %>%
  hc_xAxis(title = list(text = "")) %>% 
   hc_yAxis(title = list(text = "Counts")) %>% 
  hc_chart( options3d = list(enabled = TRUE, beta = 0, alpha = 0)) %>%  hc_title(text = "Severity")

```
